Infinite Diameter Confidence Sets in Hedges' Publication Bias Model

### 2020-08-17 infiniteci.tex
:pear: Made arXiv version.

\documentclass[article]{ajs}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{enumitem}
\setlist[enumerate]{label = (\roman*)}

%%%%%%%%%%%%
%% Macros %%
%%%%%%%%%%%%

\providecommand{\tabularnewline}{\\}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{definition}
\newtheorem{example}[thm]{\protect\examplename}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\theoremname}{Theorem}

@@

\author{Jonas Moss \\ University of Oslo}
\title{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}

\Plainauthor{Jonas Moss} 
\Plaintitle{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model} 
\Shorttitle{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}

\Abstract{
Meta-analysis, the statistical analysis of results from separate studies,
is a fundamental building block of science. But the assumptions of
classical meta-analysis models are not satisfied whenever publication
bias is present, which causes inconsistent parameter estimates. Hedges'
selection function model takes publication bias into account, but
estimating and inferring with this model is tough. Using a generalized
Gleser--Hwang theorem, we show there is no confidence set of guaranteed
finite diameter for the parameters of Hedges' selection model. This
result provides a partial explanation for why inference with Hedges'
selection model is fraught with difficulties.
}

\Keywords{meta-analysis, confidence intervals, file-drawer problem, publication bias, selection models, weight function models}
\Plainkeywords{meta-analysis, confidence intervals, file-drawer problem, publication bias, selection models, weight function models}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}
%% \setcounter{page}{1}
\Pages{1--xx}

\Address{
  Jonas Moss\\
  Department of Mathematics\\
  University of Oslo\\
  PB 1053, Blindern, NO-0316, Oslo, Norway\\
  E-mail: \email{jonasmgj@math.uio.no}
}

\begin{document}

@@

Most meta-analytic techniques assume honest and unbiased reporting
of results. But there is ample evidence that the scientific literature
is not unbiased, as studies with\emph{ }significant \textit{p}-values
tend to be published with a greater probability than other studies
\citep{Easterbrook1991-ph}, a phenomenon called publication bias
by \citet{sterling1959publication} and the file-drawer problem by
\citet{Rosenthal1979-pm}. When publication bias is present, there
is no reason to trust the results of meta-analytic methods that do
not account for it, as the parameter estimates will be inconsistent
\citep{Carter2019-rw}.

@@

Estimation of Hedges' model is difficult. \citet[Section 6.3]{Stanley2005-ng}
discusses three problematic cases in economics when maximum likelihood
was used to estimate Hedges' publication bias model. \citet[Appendix A]{McShane2016-rb}
notes that while estimation of the basic publication bias model is
hard, introducing the heterogeneity parameter exacerbates the problem.
The likelihood function has contours following approximately $\tau\propto|\theta|^{1/2}$.
Figure \ref{fig:contour lines} shows the contour lines for the meta-analysis
of \citet{Cuddy2018-kp} where the selection probabilities of the
step function (\ref{eq:step function}) are fixed at $\rho=(1,0.6,0.1)$ and $\alpha = (0, 0.025, 0.05, 1$. Only just around the maximum at $\hat{\theta}_{0}=0.55$ and $\hat{\sigma}=5\cdot10^{-7}$
can the likelihood be approximated with a quadratic function.

\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.5]{chunks/cuddy}
\par\end{centering}
\caption{\label{fig:contour lines}Contour lines for the log-likelihood for
the simple publication model using power posing data of \citet{Cuddy2018-kp}. }

\end{figure}

@@

The following example should make Definition \ref{def:confidence set}
familiar.
\begin{example}
\label{exa:t-test} Consider the usual \emph{t}-confidence interval
with $n$ observations. In this case, $\mathcal{P}$ contains all measures
$P_{\mu,\sigma}^{n}$, where $P_{\mu,\sigma}$ is the probability
measure of a normal with mean $\mu$ and standard deviation $\sigma$.
Here $Q^{n}$ denotes the $n$-fold product measure of $Q$, corresponding to $n$ independent samples from $Q$, when $Q$ is a probability measure. Let $\pi(\mu)=\{P_{\mu,\sigma}^{n}\mid\sigma>0\}$
contain all normal probability measures with mean $\mu$ and some
positive standard deviation. Then $\{\pi(\mu)\},\mu\in\mathbb{R}$
defines a partition of $\mathcal{P}$. A two-sided \emph{t}-confidence
interval is a confidence set of size $\alpha$ with partition $\{\pi(\mu)\},\mu\in\mathbb{R}$
according to Definition \ref{def:confidence set}.
\end{example}

@@

The diameter of a confidence set is the random variable
\begin{equation}
D(\omega)=\sup_{\pi\in\Pi}\{\left\Vert \pi\right\Vert \mid\omega\notin R(\pi)\}.\label{eq:diameter}
\end{equation}
That is, the diameter tells you the size of the largest accepted $\pi$.
We will assume that $D$ is Borel measurable.
\begin{defn}
\label{def:infinite diameter}A confidence set has infinite diameter
with $P$-positive probability if $P(D=\infty)>0$. It has infinite
diameter with positive probability if $P(D=\infty)>0$ for all $P\in\mathcal{P}$.
\end{defn}

This is a modified Gleser--Hwang theorem \citep[Theorem 1]{gleser1987nonexistence}
formulated in terms of partitions. Its proof is in the appendix (p.
\pageref{proof:Gleser--Hwang}).
\begin{thm}[Gleser--Hwang theorem]
\label{thm:Gleser--Hwang} Suppose there is a sequence $\{p_{n}\}$
of densities from $\mathcal{P}$ satisfying the following: 
\begin{enumerate}
\item There is a density $p^{\star}$ such that $p_{n}$ converges to $p^{\star}$
pointwise,
\item $\supp p\supseteq\supp p^{\star}$ for all densities $p$ derived
from $\mathcal{P}$,
\item the size of the equivalence class $[p_{n}]$ goes to infinity as $n$
increases, $\left\Vert [p_{n}]\right\Vert \to\infty$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{thm}

@@

Fieller's problem is the best known case of a badly behaved confidence
set. Let $X,Y$ be an observation from a bivariate normal $N([\mu_{1},\mu_{2}],I\sigma^{2})$,
where $\sigma^{2}$ is known. We want to form a confidence set for
the ratio $\mu_{2}/\mu_{1}$. The most famous confidence set is due
to \citet{Fieller1940-lg}. His confidence set can be finite, the
whole real line, or the union of two disjoint semi-infinite intervals, all with positive probability \citep{Koschat1987-dk}.

@@

\bibliography{infiniteci}

>>>>

% \documentclass[english]{article}
% \usepackage[T1]{fontenc}
% \usepackage[utf8]{inputenc}
% \usepackage[english]{babel}
% \usepackage{amsmath}
% \usepackage{amsthm}
% \usepackage{amssymb}
% \usepackage{graphicx}
% \usepackage{natbib}
% \usepackage{enumitem}
% \setlist[enumerate]{label = (\roman*)}

% %%%%%%%%%%%%
% %% Macros %%
% %%%%%%%%%%%%

% \providecommand{\tabularnewline}{\\}
% \numberwithin{equation}{section}
% \numberwithin{figure}{section}
% \theoremstyle{plain}
% \newtheorem{thm}{\protect\theoremname}
% \theoremstyle{definition}
% \newtheorem{defn}[thm]{\protect\definitionname}
% \theoremstyle{definition}
% \newtheorem{example}[thm]{\protect\examplename}
% \theoremstyle{plain}
% \newtheorem{lem}[thm]{\protect\lemmaname}
% \providecommand{\definitionname}{Definition}
% \providecommand{\examplename}{Example}
% \providecommand{\lemmaname}{Lemma}
% \providecommand{\theoremname}{Theorem}

% \renewcommand{\sqrt}[1]{{(#1)^{1/2}}}
% \DeclareMathOperator{\tr}{tr}
% \DeclareMathOperator{\Tr}{Tr}
% \DeclareMathOperator{\Var}{Var}
% \DeclareMathOperator{\argmin}{argmin}
% \DeclareMathOperator{\Cor}{Cor}
% \DeclareMathOperator{\Cov}{Cov}
% \DeclareMathOperator{\diag}{diag} 
% \DeclareMathOperator{\argmax}{argmax}
% \DeclareMathOperator{\supp}{supp}

% \author{Jonas Moss \\ University of Oslo}
% \title{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}

% \Plainauthor{Jonas Moss} 
% \Plaintitle{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model} 
% \Shorttitle{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}

% \Abstract{
% Meta-analysis, the statistical analysis of results from separate studies,
% is a fundamental building block of science. But the assumptions of
% classical meta-analysis models are not satisfied whenever publication
% bias is present, which causes inconsistent parameter estimates. Hedges'
% selection function model takes publication bias into account, but
% estimating and inferring with this model is tough. Using a generalized
% Gleser--Hwang theorem, we show there is no confidence set of guaranteed
% finite diameter for the parameters of Hedges' selection model. This
% result provides a partial explanation for why inference with Hedges'
% selection model is fraught with difficulties.
% }

% \Keywords{meta-analysis, confidence intervals, file-drawer problem, publication bias, selection models, weight function models}
% \Plainkeywords{meta-analysis, confidence intervals, file-drawer problem, publication bias, selection models, weight function models}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}
%% \setcounter{page}{1}
% \Pages{1--xx}

% \Address{
%   Jonas Moss\\
%   Department of Mathematics\\
%   University of Oslo\\
%   PB 1053, Blindern, NO-0316, Oslo, Norway\\
%   E-mail: \email{jonasmgj@math.uio.no}
% }

%% LyX 2.3.4.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,twoside]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[natbibapa]{apacite}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{definition}
\newtheorem{example}[thm]{\protect\examplename}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{infiniteci}

\titletag{An arXiv preprint, v2}
\makeatletter
\titlerunning{\@title}
\makeatother
\authorrunning{Jonas Moss}

@@

\usepackage{enumitem}
\setlist[enumerate]{label = (\roman*)}

\makeatother

\usepackage{babel}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\theoremname}{Theorem}

\begin{document}
\title{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}
\author{Jonas Moss \orcid{0000-0002-6876-6964}\\
Department of Mathematics, University of Oslo\\
PB 1053, Blindern, NO-0316, Oslo, Norway\emph{}\\
\emph{jonasmgj@math.uio.no}}
\maketitle
\begin{abstract}
Meta-analysis, the statistical analysis of results from separate studies,
is a fundamental building block of science. But the assumptions of
classical meta-analysis models are not satisfied whenever publication
bias is present, which causes inconsistent parameter estimates. Hedges'
selection function model takes publication bias into account, but
estimating and inferring with this model is tough. Using a generalized
Gleser--Hwang theorem, we show there is no confidence set of guaranteed
finite diameter for the parameters of Hedges' selection model. This
result provides a partial explanation for why inference with Hedges'
selection model is fraught with difficulties.
\end{abstract}
\keywords{meta-analysis \and confidence intervals \and file-drawer  problem \and publication bias \and selection models \and weight function models}

@@

Most meta-analytic techniques assume honest and unbiased reporting
of results. But there is ample evidence that the scientific literature
is not unbiased, as studies with significant \textit{p}-values
tend to be published with a greater probability than other studies
\citep{Easterbrook1991-ph}, a phenomenon called publication bias
by \citet{sterling1959publication} and the file-drawer problem by
\citet{Rosenthal1979-pm}. When publication bias is present, there
is no reason to trust the results of meta-analytic methods that do
not account for it, as the parameter estimates will be inconsistent
\citep{Carter2019-rw}.

@@

Estimation of Hedges' model is difficult. \citet[Section 6.3]{Stanley2005-ng}
discusses three problematic cases in economics when maximum likelihood
was used to estimate Hedges' publication bias model. \citet[Appendix A]{McShane2016-rb}
notes that while estimation of the basic publication bias model is
hard, introducing the heterogeneity parameter exacerbates the problem.
The likelihood function has contours following approximately $\tau\propto|\theta|^{1/2}$.
\begin{figure}[t]
\noindent \begin{centering}
\includegraphics[scale = 0.5]{chunks/cuddy}
\par\end{centering}
\caption{\label{fig:contour lines}Contour lines for the log-likelihood for
the simple publication model using power posing data of \citet{Cuddy2018-kp}. }
\end{figure}
Figure \ref{fig:contour lines} shows the contour lines for the meta-analysis of \citet{Cuddy2018-kp} where the selection probabilities of the step function (\ref{eq:step function}) are fixed at $\rho=(1,0.6,0.1)$ and $\alpha = (0, 0.025, 0.05, 1$. Only just around the maximum at $\hat{\theta}_{0}=0.55$ and $\hat{\sigma}=5\cdot10^{-7}$
can the likelihood be approximated with a quadratic function.

@@

The following example should make Definition \ref{def:confidence set}
familiar.
\begin{example}
\label{exa:t-test} Consider the usual \emph{t}-confidence interval
with $n$ observations. In this case, $\mathcal{P}$ contains all measures
$P_{\mu,\sigma}^{n}$, where $P_{\mu,\sigma}$ is the probability
measure of a normal with mean $\mu$ and standard deviation $\sigma$.
Here $Q^{n}$ denotes the $n$-fold product measure of $Q$, corresponding to $n$ independent samples from $Q$ when $Q$ is a probability measure. Let $\pi(\mu)=\{P_{\mu,\sigma}^{n}\mid\sigma>0\}$
contain all normal probability measures with mean $\mu$ and some
positive standard deviation. Then $\{\pi(\mu)\},\mu\in\mathbb{R}$
defines a partition of $\mathcal{P}$. A two-sided \emph{t}-confidence
interval is a confidence set of size $\alpha$ with partition $\{\pi(\mu)\},\mu\in\mathbb{R}$
according to Definition \ref{def:confidence set}.
\end{example}

@@

The diameter of a confidence set is the random variable
\begin{equation}
D(\omega)=\sup_{\pi\in\Pi}\{\left\Vert \pi\right\Vert \mid\omega\notin R(\pi)\}.\label{eq:diameter}
\end{equation}
The diameter tells you the size of the largest accepted $\pi$.
We will assume that $D$ is Borel measurable.
\begin{defn}
\label{def:infinite diameter}A confidence set has infinite diameter
with $P$-positive probability if $P(D=\infty)>0$. It has infinite
diameter with positive probability if $P(D=\infty)>0$ for all $P\in\mathcal{P}$.
\end{defn}

This is a modified Gleser--Hwang theorem \citep[Theorem 1]{gleser1987nonexistence}
formulated in terms of partitions. Its proof is in the appendix (p.
\pageref{proof:Gleser--Hwang}).
\begin{thm}[Gleser--Hwang theorem]
\label{thm:Gleser--Hwang} Suppose there is a sequence $\{p_{n}\}$
of densities derived from $\mathcal{P}$ satisfying the following: 
\begin{enumerate}
\item There is a density $p^{\star}$ such that $p_{n}$ converges to $p^{\star}$
pointwise,
\item $\supp p\supseteq\supp p^{\star}$ for all densities $p$ derived
from $\mathcal{P}$,
\item the size of the equivalence class $[p_{n}]$ goes to infinity as $n$
increases, $\left\Vert [p_{n}]\right\Vert \to\infty$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{thm}

@@

Fieller's problem is the best known case of a badly behaved confidence
set. Let $(X,Y)$ be an observation from a bivariate normal $N([\mu_{1},\mu_{2}],I\sigma^{2})$,
where $\sigma^{2}$ is known. We want to form a confidence set for
the ratio $\mu_{2}/\mu_{1}$. The most famous confidence set is due
to \citet{Fieller1940-lg}. His confidence set can be finite, the
whole real line, or the union of two disjoint semi-infinite intervals, all with positive probability \citep{Koschat1987-dk}.

@@

\bibliographystyle{apacite}
\bibliography{infiniteci}
